---
title: "MotorTrend - TransmissionType effect over MPG"
author: 'andhdo: AndresHurtado'
date: "18 de octubre de 2015"
description: Coursera's Regression Models Course, Project Assignment
output: pdf_document
keep_md: yes
geometry: margin=1cm
classoption: a4paper
---
## Executive Summary

Motor Trend, a magazine about automobile industry is interested in explore the relationshp between a set of 11 variables extracted over a collection of 32 car measures with its oil consumption (expressed in miles per gallon).
For the analysis is used the 1974's MotorTrend dataset to answer the following questions:

* "Is a transmission type (automatic or manual) better for MPG"
* "How different is the MPG betweeen both"

The conclusion is that Manual Transmission achieves a higher value of MPG compared to Automatic Transmission (near 1.8 MPG), based on a linear regression model; however this difference is statistically insignificant

## Exploratory Data Analysis
First we load the dataset (mtcar), look for information about the dataset (?mtcars) and do some preprocessing tasks. Specially we need to set cylinders (cyl) and type of transmission - automatic/manual (am) variables as categorical (factor variable).
```{r}

library(ggplot2)
data(mtcars)

mtcars$cyl  <- as.factor(mtcars$cyl)
mtcars$am   <- factor(mtcars$am,labels=c("Automatic","Manual"))
mtcars$vs   <- as.factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```

Then we explore relationships between variables (see appendix), where we can see a strong correlation between mpg and am, starting first from a visualization between all variables of the dataset (appendix 1) and second, if we go to a more detailed diagram between the variables of interest for this study (mpg and am shown in appendix 2). Now lets see the application of the model.

## Model Analysis: Using Regression
We build linear regressions over the variables of the dataset to explore it as a predictors of mpg (using AIC algorithm, that is calling lm repeatedly and finding the better selection for mpg prediction)
```{r, results="hide"}
initialmodel <- lm(mpg ~ ., data = mtcars)
summary(initialmodel) # results hidden
```
The initial model tell us that the model can explain near 78% of the variance of the mpg variable but none of the coeficients
are significant at 0.05-significant-level.
```{r, results="hide"}
bestmodel <- step(initialmodel, direction = "both")
```
see the results below
```{r}
summary(bestmodel)
```
It tell us that cyl,hp,wt cofounders and am as independent variable are candidates to fit the best model (act as predictors) and more than 84% of variability is explained combining those values as predictors (0.84 r-squared value). 
Now let's consider only the am variable and compare it with the previous model using anova:
```{r}
basemodel <- lm(mpg ~ am, data = mtcars)
anova(basemodel, bestmodel)
```
The p-value is highly significant and we reject the hypothesis that cofounder variables (cyl,hp,wt) don't contribute to the accuracy of the model.

## Model Analysis: Residuals
Residuals plots help us to find some potential problems with the model (see appendix 3). it tell us:
- residuals-vs-fitted: are distributed random, so, it verifies independence condition
- normal-q-q: (points near of the line): residuals normally distributed
- scale-location: constant variance
- increased leverage of outliers in the top right of the plots

To find points with more leverage, let's do some computations:
```{r}
leverage <- hatvalues(bestmodel)
tail(sort(leverage),3)
```
To find points that influences the model, let's do some computations:
```{r}
influential <- dfbetas(bestmodel)
tail(sort(influential[,6]),3)
```

##Model Analysis: Statistical Inference
We perform a t-test of the manual and automatic subsets. Now test the following:
  H(0): come from same distribution

```{r}
t.test(mpg ~ am, data = mtcars)
```

So we reject H0 (that is: we reject that mpg distributions for manual and automatic are the same)

## Conclusions

- mean for mpg of manual transmission is about 7 more than automatic transmission cars (24.392 - 17.147).
best model assumption tell us that:
- increasing cylinders to 6 or 8, then mpg will decrease in a factor of -3.03 or -2.16 respectively ( summary(bestmodel)$coef )
- increments in hp, causes only a decrease in 0.32 mpg (unit: 10hp)

- increments in weight, causes a decrease of -2.49 mpg (unit:1000lb)


## Appendix

### 1. Relationship between variables of the dataset 
```{r}
pairs(mpg ~ ., data = mtcars)
```
We can see higher correlations between variables like wt,disp,cyl,hp
In the quadrant mpg vs am, we can see two lines, one for automatic and one for manual. It shows variations over mpg influence.
Let's zoom it with a box plot:

### 2. Visualization of automatic vs manual transmission
```{r}
boxplot(mpg ~ am, data = mtcars, col = (c("red","blue")), ylab = "Miles Per Gallon", xlab = "Transmission Type")
```
It shows that manual transmission yields higher values of mpg

### 3. Visualization of residuals
```{r}
par(mfrow=c(2, 2))
plot(bestmodel)
```
